---
title: "FSD3201 Barometer for Swedish-speaking Finns 2009"
format: 
  html:
    self-contained: true
editor: visual
---


```{r, echo = FALSE, message=FALSE,warning=FALSE}
library(dplyr)
d <- read.csv2("daF3201_swe.csv") |> 
  select(region = q1_lk, age = q2_lk, gender = q3,
         starts_with("q16"), 
         starts_with("q17"),
         starts_with("q20"),
         starts_with("q23")
  ) |> 
  mutate(
    region = factor(
      region, levels = 1:9, 
      labels = c("Mellannyland",
                 "Östra Nyland",
                 "Västra Nyland",
                 "Åboland",
                 "Åland",
                 "Södra Österbotten",
                 "Mellersta Österbotten",
                 "Norra Österbotten",
                 "Finska Finland")),
    age = factor(
      age, levels = 1:6,
      labels = c("18-24",
                 "25-34",
                 "35-44",
                 "45-54",
                 "55-64",
                 "65-80")),
    #The media consumption has categories "No", "Sometimes", "Often", "Daily".
    consumes_swedish_radio =
      if_else(q16 == 2, q16_6 > 2, FALSE), # Sveriges Radio
    consumes_swedish_tv = 
      if_else(q17 == 2, q17_4 > 2 | q17_5 > 2, FALSE), #SVT or TV4
    consumes_swedish_news = 
      if_else(q20 == 2, q20_6 > 2 | q20_7 > 2, FALSE), #Aftonbladet/Expressen/Other
    consumes_swedish_social_media = 
      q23_8 > 2, 
  ) |> 
  select(region, age, gender, starts_with("consumes")) |> 
  filter(region != "Åland")
d <- d[rowSums(is.na(d[,4:7])) != 4, ] #remove cases where all answers are missing
# Create the aggregate Swedish media consumption variable and recode region and age:
d <- d |> 
  rowwise() |> 
  mutate(
    consumes_swedish_media = 
      any(
        consumes_swedish_radio, consumes_swedish_tv, consumes_swedish_news, 
        consumes_swedish_social_media)
  )
d <- d |> 
  mutate(
    region = case_match(
      region,
      c("Mellannyland", "Östra Nyland", "Västra Nyland") ~ "Uusimaa",
      "Åboland" ~ "Soutwest Finland",
      c("Södra Österbotten", "Mellersta Österbotten", "Norra Österbotten") ~ "Ostrobothnia",
      "Finska Finland" ~ "Other"
      
    ),
    age = case_match(
      age,
      c("25-34", "35-44") ~ "25-44",
      .default = "Others"
    ),
    gender = factor(gender, levels = 1:2, labels = c("Women", "Men"))
  )
```



The data for Barometer for Swedish-speaking Finns 2009 can be obtained from https://services.fsd.tuni.fi/catalogue/FSD3201. Here we see how the consumption of media from Sweden varies geographically in Finland. The response `uses_swedish_media` is based on questions `q16_6` (radio), `q17_4` (tv), `q20_6` (news), and `q23_8` (social media), see the descriptions in the data catalogue. If person answered "often" or "daily" to any of the questions, we coded `uses_swedish_media` as `yes`. If the person answered all questions as "rarely" or "sometimes", the answer was coded as `no`, and in other cases the answer was set to `NA`.

Results show that in Ostrobothnia, the consumption of Sweden-based media is somewhat higher than in other parts of Finland:



```{r, echo = FALSE}
knitr::kable(
  prop.table(
    table(
      region = d$region, 
      uses_swedish_media = factor(d$consumes_swedish_media, labels = c("no", "yes")), 
      useNA = "ifany"),
    1), digits = 2
)
```
By region and age:

```{r, echo = FALSE}
knitr::kable(
  prop.table(
    table(
      region = d$region, 
      uses_swedish_media = factor(d$consumes_swedish_media, labels = c("no", "yes")), 
      age = d$age,
      useNA = "ifany"),
    c(1, 3)), 
  digits = 2)
```

By region and gender:

```{r, echo = FALSE}
knitr::kable(
  prop.table(
    table(
      region = d$region, 
      uses_swedish_media = factor(d$consumes_swedish_media, labels = c("no", "yes")),
      gender = d$gender,
      useNA = "ifany"),
    c(1, 3)), 
  digits = 2
)
```
